\begin{Note}

A typical instance is
\[
\HostParameters = (L,E,d,b),
\qquad
\ParameterMap(\HostParameters)
=
\bigl(
L,
E,
\mathcal F(d,b)
\bigr),
\]
where \(\mathcal F(d,b)\) generates a fiber discretization from geometric dimensions.

The kernel differentiates the response with respect to the native parameters
\[
\CoreParameters
=
\bigl(
L,
E,
\mathcal F(d,b)
\bigr),
\]
while the host differentiates the generator \(\mathcal F\). Therefore
\[
\D \bigl(\ResponseMap_{\ModelFamily}\circ\ParameterMap\bigr)(L,E,d,b)
[\dot L,\dot E,\dot d,\dot b]
=
\D \ResponseMap_{\ModelFamily}\bigl(L,E,\mathcal F(d,b)\bigr)
\Bigl[
\D \ParameterMap(L,E,d,b)[\dot L,\dot E,\dot d,\dot b]
\Bigr].
\]
This yields derivatives with respect to \(d\) and \(b\) even though \(d\) and \(b\) are not native kernel parameters.

\end{Note}

Consider a cantilever beam
of length~$L$, loaded by a tip force~$H$, whose cross-section is a
solid rectangle of width~$b$ and depth~$d$. 

The beam is modeled with
fiber-section frame elements, so that the cross-section is
represented by a grid of $N$~integration points, each characterised by its
a location $(y_k,z_k) \in\Section$ and weight $A_k$.

% \paragraph{Core parameters.}
The quantities that enter the finite element residual directly are
\begin{equation*}
    \corey = \bigl(L,\; E,\; y_1,\, z_1,\, A_1,\;\ldots,\;
                    y_N,\, z_N,\, A_N\bigr)
    \;\in\; \mathbb{R}^{2+3N},
\end{equation*}
where $E$ is Young's modulus. 
The DDM can deliver
$\partial u/\partial y_i$ for each of these $2+3N$ parameters, where
$u$ is the tip displacement.

The analyst, however, is interested in the dependence of~$u$ on the
section \emph{geometry}:
\begin{equation}\label{eq:example-host}
    \haty = (L,\; E,\; b,\; d) \;\in\; \mathbb{R}^{4}.
\end{equation}

The parameter map~$\Ymap$ generates the fiber layout from the section
dimensions.  Given a regular $n_x \times n_y$ grid over the rectangle
$[-b/2,\,b/2]\times[-d/2,\,d/2]$, the fiber coordinates and areas
are deterministic, differentiable functions of~$b$ and~$d$:
\begin{equation*}
    (b,\,d) \;\xmapsto{\;\Ymap\;}\;
    \bigl\{(y_k,\, z_k,\, A_k)\bigr\}_{k=1}^{N}.
\end{equation*}

% \paragraph{End-to-end gradient.}
Composing $\Mmap$ with $\Ymap$ and applying the chain rule yields the
sensitivity of the tip displacement to the section depth:
\begin{equation*}
    \frac{\partial u}{\partial d}
    = \sum_{k=1}^{N}
      \biggl(
        \frac{\partial u}{\partial y_k}\,\frac{\partial y_k}{\partial d}
      + \frac{\partial u}{\partial z_k}\,\frac{\partial z_k}{\partial d}
      + \frac{\partial u}{\partial A_k}\,\frac{\partial A_k}{\partial d}
      \biggr),
\end{equation*}
where the terms $\partial u/\partial y_k$,
$\partial u/\partial z_k$, $\partial u/\partial A_k$ are provided by
the DDM inside the kernel, and the terms
$\partial y_k/\partial d$, $\partial z_k/\partial d$,
$\partial A_k/\partial d$ are computed by AD of the fiber-generation
function~$\Ymap$.

For the Euler--Bernoulli cantilever, the tip displacement under a
point load~$H$ is
\begin{equation*}
    u = \frac{H L^{3}}{3\,E\,I_y} + \frac{H\,L}{G\,A},
\end{equation*}
where $I_y = b\,d^{3}/12$ is the second moment of area and
$G = E/\bigl(2(1+\nu)\bigr)$ is the shear modulus.  
The analytical derivative with respect to~$d$ is
\begin{equation*}
% \label{eq:example-exact-deriv}
    \frac{\partial u}{\partial d}
    = \frac{\partial u}{\partial I_y}\,\frac{\partial I_y}{\partial d}
    + \frac{\partial u}{\partial A}\,\frac{\partial A}{\partial d}
    = -\frac{H L^{3}}{3\,E\,I_y^{2}}\;\frac{b\,d^{2}}{4}
      -\frac{H\,L}{G\,A^{2}}\;b,
\end{equation*}
and serves as a verification target.

The composed numerical gradient
$\partial u/\partial d$ computed through the framework agrees with
this expression to the precision of the fiber discretisation.

