Part~I develops a framework for beam section modeling that is compatible with geometrically exact kinematics, supports multiaxial material response, and recovers continuum-consistent behavior for both uniform and nonuniform warping. 

The developments of Part~I are summarized by Box~\ref{box:model-summary}.
+The essential features of the work are as follows:
+\begin{itemize}
+    \item
+\end{itemize}



\section{Implementation}

The pre-processing procedure of Box~\ref{box:prep} may be conveniently performed with a specialized software library. Notable examples include 
\begin{itemize} 
\item \texttt{section-properties} \cite{leeuwen2024sectionproperties} 
\item \texttt{xsection}
\end{itemize}

\begin{ambox}[Summary of frame models]{box:model-summary}

\RepeatEquation{eq:trace-map}

\begin{enumerate}
    % \item \textbf{Trace State}
    %     Trace Strains:
    %     \RepeatEquation{eq:def-gam-kap}
    \item Element Kinematics
        \begin{description}
            \item[$\circ$ Finite] (Box~\ref{box:exact-shear-svq})
            \RepeatEquation{eq:finite-alignment-field}
            \item[$\circ$ Linear] (Box~\ref{box:basic-shear})
            \RepeatEquation{eq:trace-lift-linear}
        \end{description}
    \item Section Kinematics 
        % \RepeatEquation{eq:saint-venant-stress}
        \RepeatEquation{eq:enhan-strain}
        with
        \RepeatEquation{eq:strain-tensor-decomp}
        \begin{description}
            \item[$\circ$ Wagner] \RepeatEquation{eq:wagner-gl-strain-pi}
            \item[$\circ$ Linear] \RepeatEquation{eq:linear-gl-strain-pi}
        \end{description}

    \item Enhanced Shapes
     \begin{description}
        \item[$\circ$ Type I: Uniform] 
          % Trace Matrix:
          % \RepeatEquation{eq:T0-inv}
          % \begin{description}
          % \item[$\circ$ Petrov] 
          \RepeatEquation{eq:uniform-shapes}
          % \item[$\circ$ Bubnov]
          % \RepeatEquation{eq:bubnov-section}
          % \end{description}
        \item[$\circ$ Type II: Nonuniform]
        \RepeatEquation{eq:nonuniform-shapes}
     \end{description}

\end{enumerate}
\end{ambox}


\subsection{Implementation}

\begin{ambox}[Section Preprocessing.]{box:prep}
\paragraph{Objective.}
Given a cross section \(\Section \subset \mathbb{R}^2\) where each point is assigned a tensorial stress-strain material \(\TensorMaterial: \StrainTensor \mapsto \StressTensor\), 
define a \emph{cross section discretization} \(\Section^h\).
\paragraph{Procedure.}
\begin{enumerate}
    \item Discretize the section \(\Section\) into quadrature points \(\{\PlanePosition_{\ell}\}\) and weights \(\{\dd A_{\ell}\}\)
    \item Solve the BVPs \cref{eq:bvp-warp-twist-iesan} for \(\WarpTwistIesan\) and \cref{eq:warp-fa-bvp} for \(\WarpFA\), and for each point \(\PlanePosition_{\ell}\) in the discretization store 
    \begin{itemize}
        \item \(\TotalStrainMatrix(\PlanePosition_{\ell})\) \eqref{eq:def-section-kinematic-matrix-Phi} or \eqref{eq:def-section-kinematic-matrix-wagner}.
        \item \(\EnhanShapes(\PlanePosition_{\ell})\) from \cref{sec:timo} or \eqref{eq:nonuniform-shapes}.
        % \item \(~~\WarpSV(\PlanePosition_{\ell}) \in \mathbb{R}\)
        % \item \(\nabla \WarpSV(\PlanePosition_{\ell}) \in \mathbb{R}^2\)
        % \item \(\nabla \WarpFA (\PlanePosition_{\ell}) \in \mathbb{R}^2\)
    \end{itemize}
\end{enumerate}
\end{ambox}

\iftrue
\begin{ambox}[Material State Determination.]{box:mati}
\paragraph{Objective.}
Given a strain vector \(\StrainVector \in \mathbb{R}^3\) and tensorial stress-strain material model \(\TensorMaterial \colon \mathrm{Sym}(\mathbb{R}^3, \mathbb{R}^3) \rightarrow \mathrm{Sym}(\mathbb{R}^3, \mathbb{R}^3)\), return the stress vector \(\StressVector \in \mathbb{R}^3\) representing the constrained stress tensor.
% \[
% \StressTensor = \StressVector \symbun \FrameBasis
% \]
\paragraph{Procedure.}
Perform Newton-Raphson iterations solving for the strain tensor \(\tilde{\StrainTensor}\) that satisfies the constraints
\[
\FrameBasis_{\alpha}\cdot  \TensorMaterial\bigl(\StrainVector \symbun \FrameBasis + \tilde{\StrainTensor} \bigr)\, \FrameBasis_{\beta} = 0
\]
Details of the procedure are developed by \cite{klinkel2002using}.
\end{ambox}
\fi 


\cite{chadha2019comprehensive} investigate the a model that accounts for transverse warping induced through axial loading of a beam.

