% \begin{Details}

% \begin{Aside}
% Beam finite element formulations have, over several decades, proliferated into a large and fragmented collection of special-case implementations.
% Production codes typically offer distinct element types for thin-walled sections, for asymmetric sections, for large displacements, for warping torsion, and so on---each developed and validated in relative isolation.
% An improvement to one formulation does not automatically benefit others; a correction to the handling of, say, shear-flexure interaction in a geometrically linear element may have no bearing on the corresponding geometrically nonlinear element, even within the same software package.
% The result is a maintenance burden that grows combinatorially with the phenomena one wishes to capture, and a validation effort that must be repeated for each combination.

% This fragmentation also obscures deficiencies.
% Certain widely-used formulations for geometrically nonlinear analysis violate strain objectivity: objective strain measures, which should be invariant under superposed rigid motions, are corrupted by the interpolation procedure \citep{crisfield1999objectivity}. 
% The errors are small for moderate rotations and vanish at converged equilibrium states, so they are often tolerated in practice.
% But tolerating such errors becomes untenable when simulations must be differentiated---as required for sensitivity analysis, gradient-based inference, or optimization---because the spurious dependence on rigid motion contaminates the computed gradients.
% Fixing the problem within existing software architectures would require rewriting element-level code, with attendant re-validation, for each affected element type.
% \end{Aside}

\section{Motivation}\label{sec:part-i-objectives}



% A family of models are developed for simulating slender members exhibiting geometric and material nonlinearities. 
Structural members are typically represented by dimensionally reduced \citep{kantorovich1958approximate} models of three dimensional continua, known as beam theories, which simplify the partial differential equations of elastostatics to produce an ordinary differential equation in one spatial variable.
When discretized by the finite element method, these models offer a substantial advantage in computational demand over direct finite element modeling of the continuum field equations, but several complications arise in the face of geometric and material nonlinear response:
\begin{description}
\item[Geometric Nonlinearity.]
Dimensional reduction of the geometrically nonlinear continuum equations generally leads to a problem posed over a constrained solution space that lacks a vectorial structure. 
Such problems lie outside the domain of classical finite element methods \citep{strang1973analysis}, which can fail to preserve important features of the constrained model, like strain objectivity and path independence \citep{jelenić1999geometrically}. 
Formulations that restore objectivity exist, but they sacrifice the symmetric tangent stiffness that accelerates iterative solution.

\item[Material Nonlinearity.]
Inelastic material response exhibits a complicated dependency on the three dimensional state of stress and strain in a body. 
Fiber section models reconstruct a uniaxial strain field by leveraging the plane section hypothesis of Euler-Bernoulli and Timoshenko beam theories. 
This approach yields acceptable accuracy when a uniaxial material model can sufficiently capture the inelastic response, as is often the case for slender members with simple cross sections not expected to experience significant torsion. 

For members subject to more general loading conditions, a multiaxial strain state must be accounted for in the resultant formulation. 
The plane section hypothesis alone fails to capture cross-sectional warping, and neglecting these effects tends to over-predict strength and stiffness. 
Depending on the problem, warping may be incorporated into a beam theory either implicitly, through appropriately constructed resultant relations, or explicitly, through auxiliary kinematic fields.

% Fiber section models \citep{taucer1991fiber} reconstruct a \emph{uniaxial} strain field by leveraging the plane section hypothesis of Euler-Bernoulli and Timoshenko beam theories. 
% This approach yields acceptable accuracy when a uniaxial material model can sufficiently capture the inelastic material response. 
% This is often the case for slender members with simple cross sections that are not expected to experience torsion. 

% It is possible to reconstruct a \emph{multiaxial} strain field from the fields of a beam theory by leveraging the plane section hypothesis, but this fails to account for cross-sectional warping and tends to over-predict strength and stiffness. 

% A more accurate approach for reconstructing a multiaxial strain state introduces auxiliary kinematic fields in the formulation of the beam theory that model this warping \citep{reissner1952nonuniform,klinkel2003anisotropic,lecorvec2012nonlinear}. 
% When no limit is placed on the number of warping fields, it is possible to represent a continuum to any desired accuracy. 
% With only one auxiliary warping field, \cite{armero2021modeling} has demonstrated that such models furnish stress fields which violate equilibrium and may produce inaccurate results. 
% However, excessive auxiliary warping fields can be computationally prohibitive. 
% Furthermore such models are often cumbersome, and it is preferable to forgo auxiliary warping fields altogether whenever possible. 

\end{description}

% Objective: Multiaxial inelastic material response in beams.
A beam theory is only able of representing a subset of motions which a body may experience. Increasing the number of warping fields enlarges this class, but it also increases the number of section-level unknowns and the associated boundary conditions and response quantities that must be specified and interpreted. 
It is therefore of considerable practical interest to determine, for a given kinematic description, the most faithful representation of the multiaxial strain state that is accessible without further enrichment. 

This work proposes an enhanced strain principle for beam sections that improves the three-dimensional strain field used for multiaxial constitutive integration without altering the element kinematics, thereby maintaining compatibility with a broad class of beam finite elements. 
For a prescribed kinematic description, the strain reconstruction is augmented by section-level fields whose amplitudes are determined by an assumed-strain construction. 
The general framework is specialized for two classes of problems. 
For standard beam elements without auxiliary warping fields, the resulting section models support multiaxial inelastic response and, in the elastic limit, reproduce Saint~Venant's continuum solutions for the corresponding loading classes. 
When nonuniform warping requires auxiliary fields and continuum benchmarks are generally unavailable, the same framework enforces pointwise equilibrium in the elastic regime and improves the accuracy obtained for a prescribed set of warping modes, thereby reducing the number of auxiliary fields needed in practice.



\section{Literature}

Beam theories which are capable of representing multiaxial nonlinear material response fall into two broad categories:
\begin{itemize}
    \item Type I theories employ only the classical deflection and rotation fields as unknowns and furnish three-dimensional finite element implementations with six degrees of freedom per node.
    \item Type II theories include at least one auxiliary field to model cross sectional warping and furnish three-dimensional finite element implementations with more than six degrees of freedom per node.
\end{itemize}


\subsection{Type I}
% For uniform warping, the central open problem is handling complications associated with shear (Saint Venant's P_II).
When cross sectional warping is expected to vary uniformly over the body, auxiliary warping fields can be eliminated by associating various warping modes with classical beam deformations like curvature \(\bm{\kappa}\). 
The phenomenon of torsional warping is particularly amenable to this approach and has been successfully incorporated into beam finite elements by several authors %\citep{gruttmann1998geometrical}
, but warping induced by transverse shear introduces several difficulties:
\begin{enumerate}
    \item Warping is not trivially proportional to a classical beam deformation field
    \item The shear warping shapes arising in Saint-Venant's solution do not exhibit the same orthogonality properties as the torsion warping function.
    \item Shear-induced warping is coupled to Poisson's effect.
\end{enumerate}
In the two-dimensional setting, (2) is avoided and a satisfactory remedy for (1) may be obtained by appealing to classical shear-correction arguments.
An example of such a treatment is the investigation by \cite{simo1982bifurcation} which employed Cowper's construction \citep{cowper1966shear} to produce a two-dimensional beam theory of elastomeric bearings. 
However, as elaborated in \cref{sec:timo}, this approach cannot be made variationally consistent in the three-dimensional setting. 
A similar two-dimensional approach was taken by \cite{saritas2006mixed} to formulate a beam theory of shear-yielding structural steel members where Poisson's ratio was assumed to vanish.

In the case of pure bending, it is standard that inelastic elements maintain exact consistency with Saint-Venant's pure bending solution \citep[Art.87]{love1944treatise}, which finds bending resistance in proportion to the elastic modulus \(E\) and the second-area-tensor of the cross section \(\bm{I}\) \citep{spacone1992beam,simo1984numerical}. 
Similar consistency in torsion and elongation is also standard \citep{gruttmann1998geometrical}. 
A comparably consistent treatment for shear appears to remain an open problem.


\subsection{Type II}
% For nonuniform warping, 
When cross sectional warping is \emph{nonuniform}, auxiliary warping fields become necessary and open problems exist for both shear and torsion. 
The central problem is minimizing the number of warping shapes. 
These additional degrees 
of freedom inflate the global system of equations, incurring a cost 
that scales as the cube of the DOF count for direct linear solvers.

A pioneering work is that of \cite{timoshenko1905torsional,vlasov1961thinwalled}\footnote{A comprehensive review of nonuniform torsion is provided by \cite{armero2021modeling}}, where warping is constrained to the rate of twist to produce a higher order governing system of beam differential equations. 
This falls outside the Type I/Type II classification employed in this work as warping is not truly modeled by an independent field, but due to the higher-order governing differential equation the model cannot be realized in the classical 6-DOF/node finite element setting of Type I theories. 

\paragraph{Single-Field Twist Theories.}
An early Type II formulation is the RBV\footnote{An acronym of the original author's names coined by \cite{armero2021modeling}} model due to \cite{reissner1952nonuniform,benscoter1954theory} which incorporates a single auxiliary warping field \(\alpha\) associated to Saint~Venant's torsion warping function \(\WarpTwistIesan\). 
The unconstrained nature of this theory as compared to the Vlasov model makes it more amenable to implementation in a finite element setting. 
A notable example is the geometrically exact variant put forth by \cite{simo1991geometricallyexact} and extended by \cite{gruttmann2000theory} to investigate inelastic material response. 
This theory has also been investigated by \cite{rong2020geometrically} in the context of isogeometric analysis. 
% However, in this work the reference curve is restricted to intersect sections at the centroid. 
However, \cite{armero2021modeling} has observed that in the linear-elastic regime, the three dimensional stresses arising from this model fail to satisfy the continuum equilibrium equations, and the predicted torsional response can be inaccurate for short shafts and closed thin-walled sections. 

\paragraph{Three-Field Twist/Shear Theories.}
The foundation laid by the RBV model has been directly extended to incorporate Saint~Venant's shear warping functions in a Type II theory with three auxiliary warping fields. 
\cite{klinkel2003anisotropic} investigated this approach in the context of geometrically exact beam theory. \cite{elfatmi2007nonuniform} investigated this approach extensively in the context of linear elasticity with Poisson's effect neglected. Similar generalizations were also considered by \cite{hjelmstad1987warping,kim2005exact}. 
% See also \cite{ladeveze1998new}.

\paragraph{General Field Theories.}
\cite{lecorvec2011enhanced} % lecorvec2012nonlinear
generalized this line of work further by considering an arbitrary number of warping fields with warping shapes represented by Lagrange interpolation polynomials in a geometrically linear inelastic finite element formulation. \cite{lecorvec2012nonlinear} leveraged a corotational formulation to accurately model finite deformation benchmark problems. 
The resulting formulation is particularly powerful when several auxiliary fields are included, and has been championed by several authors \citep{dire20173d,addessi2021enriched}. 
However, these interpolation schemes had to be constructed uniquely for a given section geometry.%, and were not able to reproduce classical continuum solutions in the elastic range. 
Furthermore, due to the general nature of the Lagrange interpolation basis, several warping fields were required to accurately reproduce classical solutions. 

\paragraph{Alternative Theories.}
\citet{wackerfuss2009mixed} present a geometrically exact warping element without auxiliary fields, where warping shapes are resolved internally and do not need to be supplied at the pre-processing stage. However, the work is restricted to beams of rectangular cross sections.

\subsection{Scope and Methodology}

An assumed strain framework is presented that enhances section constitutive models for standard beam elements in Type I and II categories, and supports integrating inelastic multiaxial material response over the cross section. 


% Beam theories are typically derived for finite element analysis by one of two approaches \citep{antman1995nonlinear}:
% \begin{itemize}
% \item In an \emph{induced} approach , the field equations of continuum mechanics are stated weakly in the form of a functional\footnote{Typical functionals include virtual work, Hu-Washizu, and Hellinger-Reissner.} 
% and the displacement field is postulated. 
% \item In an \emph{intrinsic} approach, a reduced kinematic description and strain measures are postulated on a one-dimensional domain and governing equations are derived by enforcing work conjugacy.
% \end{itemize}


\begin{remark}\label{rem:inelastic-consistency}
Although the objective of this work is enhanced simulation in the nonlinear regime, the methodology of \cref{sec:timo} and \cref{sec:nonuniform} which targets linear elastic consistency is naturally a necessary feature of an inelastic model, and follows a standard protocol:
\begin{itemize}
    \item The enhanced strain fields employed for inelastic plane finite elements \citep{simo1990class} are derived from those exhibited in elastic bending \citep{wilson1990use}.
    \item The plane-section strain profile employed by standard fiber beam elements \citep{spacone1992beam} ensures agreement with the elastic solution for bending  \citep[Art. 87]{love1944treatise}.
    \item Saint Venant's linear elastic torsional warping function is employed in the geometrically nonlinear beam of \citep{simo1991geometricallyexact} and the inelastic beam of \citep{gruttmann2000theory}.
\end{itemize}
\end{remark}

\begin{remark}
By admitting an arbitrary number of auxiliary warping modes, the general beam theory outlined in \cref{sec:genwarp} generalizes the scope of classical plane-section beam theories, whose correspondence with continuum mechanics is only achieved asymptotically with the slenderness of a member \citep{parker1979asymptotic,marigo2006hierarchy,mielke1988saintvenant}. %cimetiere1988asymptotic
Each auxiliary warping field enriches the function space of the approximate model in a manner similar to mesh refinement of a continuum finite element model.
With few judiciously selected auxiliary warping fields, the beam theory of \cref{sec:genwarp} can achieve the same accuracy of a continuum finite element model with a significant reduction in computational demand. 
\end{remark}


\subsection{Outline}

% \Cref{sec:genwarp} develops equilibrium equations for a family of beam theories that account for cross-sectional warping in the geometrically exact context.
% Beam theories are classified into four types according to how warping amplitudes are related to the classical beam strains. 
% The Type~3 formulation, in which warping amplitudes are algebraically constrained to the beam strains, is of particular interest as it requires only six degrees of freedom.
\begin{description}
\item[Preliminaries.]
\Cref{sec:genwarp} introduces the general mathematical setting. 
Notation is introduced and a family of geometrically exact Cosserat beam theories with auxiliary warping is outlined. Two sources of geometric nonlinearities are identified.
The distinction between uniform and nonuniform warping is introduced and discussed. 


\item[Enhanced Strain Framework.]
\Cref{sec:enhan-section} presents a novel enhanced strain principle for formulating constitutive equations in the setting of \cref{sec:genwarp} which incorporates assumed strain shapes \(\EnhanShapes\) into the section response. General conditions are imposed on the assumed fields \(\EnhanShapes\), but specific identifications are not made until the following chapters. A general section state determination algorithm is presented. 
% This general principle is specialized in \Cref{sec:timo} and \Cref{sec:nonuniform} for uniform and nonuniform warping, respectively. 
% In both cases, the enhanced shapes are derived to ensure continuum-consistent response in the elastic regime (\cref{rem:inelastic-consistency}). 

\item[Enhanced Uniform Warping.]
\Cref{sec:timo} develops enhanced shapes \(\EnhanShapes\) that are suitable when warping is expected to be uniform. By systematically imposing agreement with Saint Venant's linear elastic continuum solutions, the space of arbitrary shape functions \(\EnhanShapes\) is reduced to a family of formulations with 16 free parameters. Within the framework of the chapter, these parameters may be selected to represent physical details like the nature of boundary conditions. \Cref{sec:trace-cowper} and \cref{sec:trace-energy} present two alternative identifications of these parameters with explicit formulae that depend on the geometry and composition of the member. 
Example~\ref{sec:frame-0012} demonstrates continuum consistency of the resulting section model using standard beam elements without auxiliary warping degrees of freedom.

\item[Enhanced Nonuniform Warping.]
\Cref{sec:nonuniform} develops enhanced shapes \(\EnhanShapes\) that are suitable when non-uniform torsional warping is expected to be significant.
In contrast with \cref{sec:timo}, continuum solutions are not available in this setting, and consistency is instead achieved by requiring pointwise equilibrium in the elastic regime. The space of shape functions is reduced to a one-parameter family. 
\end{description}




%through the reduced Type~3 models of \cref{sec:genwarp}. 
% The investigation gives rise to the 

% \end{Details}

