

\subsection{Class II: Equilibrium Consistency}\label{def:class-ii}
The traces that fall under Class I are capable of exactly representing Saint~Venant's solutions in a standard finite element context, but without further restriction they may still exhibit certain undesirable properties.
In particular, [TODO: elaborate]
\begin{itemize}
    \item Tangent symmetry
    \item Nodal loads
\end{itemize}

The conjugate stiffness is constructed from the material strain energy:
\begin{equation}
\mathcal{U}(x) \coloneq \tfrac{1}{2}\int_\Section \bm{\varepsilon}\cdot \mathbb{C}\, \bm{\varepsilon}\dA 
= \tfrac{1}{2}\bm{e}\cdot \bar{\mathbf{C}}_T(x)\,\bm{e},
\label{eq:energy-conjugacy-def}
\end{equation}


% \begin{definition}[Class II]\label{def:class-ii}
A Class I trace \(\mathcal{T}\) is a \textbf{Class II} trace if the energy-conjugate resultants $\bar{\bm{s}}$ satisfy the same equilibrium as the classical resultants \eqref{eq:equilibrium-ode}:
\begin{equation}
\bar{\bm{s}}' = \mathbb{J}\,\bar{\bm{s}}.
\label{eq:sbar-equilibrium}
\end{equation}
% \end{definition}


\begin{remark}
Prismatic $\mathbf{C}_T$ implies prismatic $\bar{\mathbf{C}}_T$ but the converse does not hold. 
Prismatic $\bar{\mathbf{C}}_T$ with non-prismatic 
$\mathbf{C}_T$ requires $\TraceEvolveEP - \IesanEvolveSP \neq \mathbf{0}$ 
but $\IesanEnergy$-skew. This is an exceptional situation. 
\end{remark}

% \begin{proposition}[S2. Equilibrium-Conjugate]
% \label{prop:equilibrium-compatibility}
Let $\mathbf{C}_T$ and $\bar{\mathbf{C}}_T$ be prismatic (C2). 
The condition 
\eqref{eq:sbar-equilibrium} holds if and only if
\begin{equation}
[\bar{\mathbf{G}}(x), \mathbb{J}] = \mathbf{0}, \qquad 
\bar{\mathbf{G}} \coloneq \bar{\mathbf{C}}_T(x)\,\mathbf{C}_T^{-1}(x)
\qquad\forall x.
\label{eq:commutator-condition}
\end{equation}
% \end{proposition}

% \begin{proof}
With prismatic stiffnesses, $\bar{\bm{s}} = \bar{\mathbf{C}}\,\bm{e}$ and 
$\snm = \mathbf{C}_T\,\bm{e}$, so $\bar{\bm{s}} = \bar{\mathbf{G}}\,\snm$. 
From \eqref{eq:equilibrium-ode}:
\begin{equation*}
\bar{\bm{s}}' = \bar{\mathbf{G}}\,\snm' = \bar{\mathbf{G}}\,\mathbb{J}\,\snm 
= \bar{\mathbf{G}}\,\mathbb{J}\,\bar{\mathbf{G}}^{-1}\,\bar{\bm{s}}.
\end{equation*}
For this to equal $\mathbb{J}\,\bar{\bm{s}}$, we require 
$\bar{\mathbf{G}}\,\mathbb{J}\,\bar{\mathbf{G}}^{-1} = \mathbb{J}$, i.e., 
$[\bar{\mathbf{G}}, \mathbb{J}] = \mathbf{0}$. 
% \end{proof}
%
The set of all matrices satisfying \eqref{eq:commutator-condition} comprise the equilibrium invariance group $G$:
\[
G \coloneq \bigl\{\,\mathbf{K} \in \mathrm{GL}(6) \;\big|\; [\mathbf{K},\mathbb{J}] = 0\,\bigr\}.
\]
This further restricts the space of trace matrices \(\TraceDeDpRoot\) to a nonlinear sub-manifold, for which a clean structure like that obtained under \eqref{eq:req-p1} does not exist.
However, it is possible to conveniently parameterize a subset of these traces, by first identifying a reference trace \(\mathcal{T}_R\) in the class. 


\subsection{Class III: Variational Consistency}\label{def:class-iii}
We now consider a final restriction which eliminates all remaining free parameters, and furnishes a unique trace matrix: the \emph{energetic} trace. 

% \begin{definition}[Class III]\label{def:class-iii}
A Class I trace is a \textbf{Class III} trace if the conjugate resultants \(\TraceTotalEnergyStress\) are exactly the classical rigid-section resultants \(\snm\):
\begin{equation}
    \bar{\bm{s}} \equiv \snm .
    \label{eq:def-s3}
\end{equation}
% \end{definition}

% \begin{proposition}[Class III]
All Class III traces are equivalent in the Saint~Venant class, with trace matrix determined by the consistency condition:
\begin{equation}
\TraceDeDpRoot = \IesanStiff^{-\top} \IesanEnergy
\label{eq:T0-formula}
\end{equation}

% \end{proposition}
% \begin{proof}
For a Class I trace with $\bm{e}(x) = \TraceDeDpRoot \exp(x\IesanEvolveSP)\bm{p}$, 
the energy tangent \eqref{eq:energy-conjugacy-def} becomes:
\begin{equation*}
\delta\bm{p}\cdot \left(\mathbf{1} + x\TraceEvolveEP^\top\right) \TraceDeDpRoot^\top \bm{s}(x) 
= \delta\bm{p}\cdot \left(\mathbf{1} + x\TraceEvolveEP^\top\right) \IesanEnergy\, \left(\mathbf{1} + x\TraceEvolveEP\right) \bm{p}.
\end{equation*}
Since this must hold for all $\bm{p}$, the class requires:
\begin{equation*}
\TraceDeDpRoot^\top \mathbf{R}(x) = \IesanEnergy\, \left(\mathbf{1} + x\TraceEvolveEP\right).
% \label{eq:conjugacy-condition}
\end{equation*}
with \(\IesanEnergy\) furnished by \eqref{eq:def-iesan-energy}. 
Expanding both sides using $\left(\mathbf{1} + x\TraceEvolveEP\right) = \mathbf{1} + x\IesanEvolveSP$ and 
matching powers of $x$:
\begin{align*}
x^0: \qquad & \TraceDeDpRoot^\top \IesanStiff = \IesanEnergy, 
\\ %\label{eq:x0-condition} \\[4pt]
x^1: \qquad & \TraceDeDpRoot^\top \mathbf{R}_1 = \IesanEnergy\,\IesanEvolveSP.
% \label{eq:x1-condition}
\end{align*}
Consistency in \(x^0\) requires:
\begin{equation*}
\TraceDeDpRoot = \IesanStiff^{-\top} \IesanEnergy
% \label{eq:T0-formula}
\end{equation*}
The linear evolution \(x^1\) yields the consistency condition:
\begin{equation}
\IesanEnergy\,\IesanStiff^{-1} \mathbf{R}_1 = \IesanEnergy\,\IesanEvolveSP
\qquad \Longrightarrow \qquad 
\IesanStiff^{-1} \mathbf{R}_1 = \IesanEvolveSP
\label{eq:consistency-condition}
\end{equation}
% \end{proof}
