
\subsection{The Adjoint (Reverse‐Mode) Method} \label{sec:adjoint}

For many-parameter optimisation one usually seeks the gradient of a scalar
objective $\Objective(\bm{u},\bm{y})$. 
Using the Lagrangian
%
\[
  \mathcal{L}(\bm{u},\bm{y},\bm{\eta})
  \;=\;
  J(\bm{u},\bm{y})
  +\bigl\langle\bm{\eta},\bm{g}(\bm{u},\bm{y})\bigr\rangle,
\]
%
stationarity with respect to $\bm{u}$ at the solution
$\bm{u}=\bm{u}(\bm{y})$ gives the \emph{adjoint equation}
%
\begin{equation}\label{eq:adjoint}
  \bigl(\partial_{\bm{u}}\bm{g}\bigr)^{\!\!\mathsf{T}}\bm{\eta}
  \;=\;
  \bigl(\partial_{\bm{u}} \Objective \bigr)^{\!\!\mathsf{T}} .
\end{equation}
or
\begin{equation}\label{eq:adjoint}
\bigl(\tfrac{\partial\bm{g}}{\partial\bm{u}}\bigr)^{\!\mathsf{T}}
\,\bm{\AdjointState}
\;=\;
\left(\frac{\partial \Objective}{\partial\bm{u}}\right)^{\!\mathsf{T}} .
\end{equation}
%
Once $\bm{\eta}$ is known, the gradient
$\nabla_{\!\bm{y}} \Objective$ follows from the \emph{vector–Jacobian product}
%
\begin{equation}\label{eq:vjp}
  \nabla_{\!\bm{y}} \Objective
  \;=\;
  -\,\bigl\langle\bm{\eta},\partial_{\bm{y}}\bm{g}\bigr\rangle
  +\partial_{\bm{y}} \Objective
\end{equation}
or, equivalently,
\begin{equation}\label{eq:grad}
\nabla_{\!\bm{y}} \Objective
\;=\;
-\bm{\AdjointState}^{\mathsf{T}}
\frac{\partial\bm{g}}{\partial\bm{y}}
\;+\;
\frac{\partial \Objective}{\partial\bm{y}} .
\end{equation}

\paragraph{Reverse-mode AD.}
The operation
$\bm{\eta}^{\mathsf{T}}\!\partial\bm{h}(\bm{y})$ in~\eqref{eq:vjp} is the
\emph{vector–Jacobian product} (VJP), the fundamental linear map produced by
\textbf{reverse-mode automatic differentiation}.
Reverse mode traverses the computational graph \emph{backwards}, reusing the
factorisation (or iterations) already computed for the primal Newton solve.

\paragraph{Complexity.}
The FLOP cost of solving~\eqref{eq:adjoint} is comparable to one primal
linearisation and is \emph{independent of $n_p$}.  The entire gradient is
available after a single VJP.  The trade-off is memory: all primal
intermediates must be stored (or recomputed via checkpointing), so memory
scales with the depth of the solve.

\paragraph{Automatic–differentiation view.}
Eqs.~\eqref{eq:adjoint}–\eqref{eq:grad} constitute \textbf{reverse–mode
automatic differentiation}: one first executes the \emph{primal} solve to
obtain $\bm{u}$, then runs a reverse sweep that back–propagates the covector
$\bm{\AdjointState}$ through every operation (row–wise evaluation of
$\partial\bm{h}^\mathsf{T}$).  The FLOP cost scales with the
\emph{number of outputs} $n_g$ (often one),
while memory grows with the depth of the computation because intermediate
states must be available for the reverse sweep \cite{MITAD2023}.

\paragraph{When is it efficient?}
Reverse mode shines when $n_g\!\ll\!n_p$ (wide Jacobians), e.g.
\begin{itemize}
  \item gradient‐based optimization of a scalar loss depending on
        millions of design variables,
  \item machine‐learning training where $\bm{y}$ are network weights.
\end{itemize}



The adjoint state space is chosen to simplify the physical
interpretation of equation
\href{constraint_(mathematics)}{constraints}.\footnote{Plessix, R-E. "A
  review of the adjoint-state method for computing the gradient of a
  functional with geophysical applications." Geophysical Journal
  International, 2006, 167(2): 495-503.
  \href{https://academic.oup.com/gji/article/167/2/495/559970}{free
  access on GJI website}}

Adjoint state techniques allow the use of
\href{integration_by_parts}{integration by parts}, resulting in a form
which explicitly contains the physically interesting quantity. An
adjoint state equation is introduced, including a new unknown variable.

The adjoint method formulates the gradient of a function towards its
parameters in a constraint optimization form. By using the dual form of
this constraint optimization problem, it can be used to calculate the
gradient very fast. A nice property is that the number of computations
is independent of the number of parameters for which you want the
gradient. The adjoint method is derived from the
\emph{dual problem} and is used e.g. in the Landweber iteration method.

When the initial problem consists of calculating the product \(s^T x\)
and \(x\) must satisfy \(Ax=b\), the dual problem can be realized as
calculating the product , where \(r\) must satisfy \(A^* r = s\). And
\(r\) is called the adjoint state vector.

\subsubsection{General case}\label{general_case}

The original adjoint calculation method goes back to \cite{cea1986conception} with the use of the Lagrangian of the optimization
problem to compute the derivative of a functional with respect to a
\href{Shape_optimization}{shape} parameter.

For a state variable \(\TangentState \in \mathcal{U}\), an optimization variable
\(v\in\mathcal{V}\), an objective functional
\(J:\mathcal{U}\times\mathcal{V}\to\mathbb{R}\) is defined. The state
variable \(u\) is often implicitly dependent on \(v\) through the
(direct) state equation \(D_{\bm{y}}(u)=0\) (usually the weak form of a partial differential equation),
thus the considered objective is \(j(v) = J(u_{\bm{y}},v)\), where \(u_{\bm{y}}\) is
the solution of the state equation given the optimization variables
\(v\). Usually, one would be interested in calculating \(\nabla j(v)\)
using the \href{chain_rule}{chain rule}:

\[\nabla j(v) = \nabla_{\bm{y}} \Objective(u_{\bm{y}},v) + \nabla_u \Objective(u_{\bm{y}})\nabla_{\bm{y}} u_{\bm{y}}.\]

Generally the term \(\nabla_{\bm{y}} u_{\bm{y}}\) is often very hard to
differentiate analytically since the dependance is defined through an
implicit equation. The Lagrangian functional can be used as a workaround
for this issue. Since the state equation can be considered as a
constraint in the minimization of \(j\), the problem

\[\text{minimize}\ j(v) = J(u_{\bm{y}},v)\]

\[\text{subject to}\ \Residual_{\bm{y}}(\TangentState_{\bm{y}}) = 0\]

has an associate Lagrangian functional
\(\mathcal{L}:\mathcal{U}\times\mathcal{V}\times\mathcal{U} \to \mathbb{R}\)
defined by

\[\mathcal{L}(u,v,\AdjointState) = J(u,v) + \langle D_{\bm{y}}(u),\AdjointState\rangle,\]

where \(\AdjointState \in\mathcal{U}\) is a Lagrange multiplier or \textbf{adjoint state variable} and \(\langle\cdot,\cdot\rangle\) is an inner product
on \(\mathcal{U}\). 
The method of Lagrange multipliers states that a
solution to the problem has to be a \href{stationary_point}{stationary
point} of the lagrangian, namely

\[\begin{cases}
d_u\mathcal{L}(u,v,\AdjointState;\delta_u) = d_uJ(u,v;\delta_u) + \langle \delta_u,D_{\bm{y}}^*(\AdjointState)\rangle = 0 & \forall \delta_u \in \mathcal{U},\\
d_{\bm{y}}\mathcal{L}(u,v,\AdjointState;\delta_{\bm{y}}) = d_{\bm{y}}J(u,v;\delta_{\bm{y}}) + \langle d_{\bm{y}}D_{\bm{y}}(u;\delta_{\bm{y}}),\AdjointState\rangle = 0 & \forall \delta_{\bm{y}}\in\mathcal{V},\\
d_\AdjointState\mathcal{L}(u,v,\AdjointState;\delta_\AdjointState) = \langle D_{\bm{y}}(u), \delta_\AdjointState\rangle = 0 \quad & \forall \delta_\AdjointState \in \mathcal{U},
\end{cases}\] where \(d_xF(x;\delta_x)\) is the
\href{Gateaux_derivative}{Gateaux derivative} of \(F\) with respect to
\(x\) in the direction \(\delta_x\). The last equation is equivalent to
\(D_{\bm{y}}(u) = 0\), the state equation, to which the solution is \(u_{\bm{y}}\).
The first equation is the so-called adjoint state equation,

\[
\langle \delta_u,D_{\bm{y}}^*(\AdjointState)\rangle = -d_uJ(u_{\bm{y}},v;\delta_u) \quad \forall \delta_u \in \mathcal{U},
\]
%
because the operator involved is the adjoint operator of \(D_{\bm{y}}\),
\(D_{\bm{y}}^*\). Resolving this equation yields the adjoint state
\(\AdjointState_{\bm{y}}\). The gradient of the quantity of interest \(j\) with
respect to \(v\) is
\(\langle \nabla j(v),\delta_{\bm{y}}\rangle = d_{\bm{y}} j(v;\delta_{\bm{y}}) = d_{\bm{y}} \mathcal{L}(u_{\bm{y}},v,\AdjointState_{\bm{y}};\delta_{\bm{y}})\)
(the second equation with \(u=u_{\bm{y}}\) and \(\AdjointState=\AdjointState_{\bm{y}}\)), thus it
can be easily identified by subsequently resolving the direct and
adjoint state equations. The process is even simpler when the operator
\(D_{\bm{y}}\) is \href{Self-adjoint_operator}{self-adjoint} or symmetric since
the direct and adjoint state equations differ only by their right-hand
side.

\subsubsection{Example: Linear case}\label{example_linear_case}

In a real finite dimensional \href{linear_programming}{linear
programming} context, the objective function could be
\(J(u,v) = \langle Au , v \rangle\), for \(v\in\mathbb{R}^n\),
\(u\in\mathbb{R}^m\) and \(A \in \mathbb{R}^{n\times m}\), and let the
state equation be \(K_{\bm{y}} u = b\), with \(K_{\bm{y}} \in \mathbb{R}^{m\times m}\)
and \(b\in \mathbb{R}^m\).

The Lagrangian function of the problem is
\(\mathcal{L}(u,v,\AdjointState) = \langle A u,v\rangle + \left< K_{\bm{y}}u - b,\AdjointState\right>\),
where \(\AdjointState\in\mathbb{R}^m\).

The derivative of \(\mathcal{L}\) with respect to \(\AdjointState\) yields the
state equation as shown before, and the state variable is
\(u_{\bm{y}} = B_{\bm{y}}^{-1}b\). The derivative of \(\mathcal{L}\) with respect to
\(u\) is equivalent to the adjoint equation, which is, for every
\(\delta_u \in \mathbb{R}^m\),

\[d_u[\langle K_{\bm{y}}\cdot- b, \AdjointState\rangle](u;\delta_u)= - \langle A^\top v,\delta u\rangle \iff \langle K_{\bm{y}} \delta_u,\AdjointState\rangle = - \langle A^\top v,\delta u\rangle \iff \langle K_{\bm{y}}^\top \AdjointState +  A^\top v,\delta_u\rangle = 0 \iff K_{\bm{y}}^\top \AdjointState = - A^\top v.\]
Thus, we can write symbolically \(\AdjointState_{\bm{y}} = K_{\bm{y}}^{-\top}A^\top v\). 
The gradient would be

\[
\langle \nabla j(v),\delta_{\bm{y}}\rangle = \langle Au_{\bm{y}},\delta_{\bm{y}}\rangle + \langle \nabla_{\bm{y}} B_{\bm{y}} : \AdjointState_{\bm{y}}\otimes u_{\bm{y}},\delta_{\bm{y}}\rangle,
\]
where \(\nabla_{\bm{y}} B_{\bm{y}} = \frac{\partial B_{ij}}{\partial v_k}\) is a
third-order \url{tensor}, \(\AdjointState_{\bm{y}} \otimes u_{\bm{y}} = \AdjointState^\top_{\bm{y}} u_{\bm{y}}\)
is the dyadic product between the direct and
adjoint states and \(:\) denotes a double tensor contraction. 
It is assumed that
\(B_{\bm{y}}\) has a known analytic expression that can be differentiated
easily.

\paragraph{Numerical consideration for the self-adjoint
case}\label{numerical_consideration_for_the_self_adjoint_case}

If the operator \(B_{\bm{y}}\) was self-adjoint, \(B_{\bm{y}} = B_{\bm{y}}^\top\), the direct
state equation and the adjoint state equation would have the same
left-hand side. In the goal of never inverting a matrix, which is a very
slow process numerically, a \href{LU_decomposition}{LU decomposition}
can be used instead to solve the state equation, in \(O(m^3)\)
operations for the decomposition and \(O(m^2)\) operations for the
resolution. That same decomposition can then be used to solve the
adjoint state equation in only \(O(m^2)\) operations since the matrices
are the same.

\subsubsection{External links}\label{external_links}

\begin{itemize}
\tightlist
\item
  A well written explanation by Errico:
  \href{https://dx.doi.org/10.1175/1520-0477(1997)078\%3C2577:WIAAM\%3E2.0.CO;2}{What
  is an adjoint Model?}
\item
  Another well written explanation with worked examples, written by
  Bradley \href{http://cs.stanford.edu/~ambrad/adjoint_tutorial.pdf}{1}
\item
  More technical explanation: A
  \href{https://doi.org/10.1111/j.1365-246X.2006.02978.x}{review} of the
  adjoint-state method for computing the gradient of a functional with
  geophysical applications
\item
  MIT course
  \href{https://web.archive.org/web/20150906095132/http://ocw.mit.edu/courses/mathematics/18-325-topics-in-applied-mathematics-waves-and-imaging-fall-2012/lecture-notes/MIT18_325F12_Chapter4.pdf}{2}
\item
  MIT notes \href{http://math.mit.edu/~stevenj/18.336/adjoint.pdf}{3}
\end{itemize}


