
\section{Discussion}
\label{sec:discussion}

The framework rests on a separation of responsibilities between
four layers:
\begin{itemize}
\item 
% \paragraph{Kernel primitive.}
The finite element library exposes the model map~$\Mmap$ as a
\emph{differentiable primitive}---an atomic operation whose primal
evaluation, JVP rule, and VJP rule are each implemented internally
using the DDM.  
From the perspective of the \autodiff{} framework, $\Mmap$ is
a black box with known first-order differential structure. 

\item
% \paragraph{Host parameter map.}
The user writes the parameter map~$\Ymap$ in the host language as
ordinary, unrestricted code.  
No special annotation or manual
derivative is required; the \autodiff{} framework differentiates~$\Ymap$
automatically.

\item The binding layer...

\item
% \paragraph{Automatic composition.}
The \autodiff{} framework composes the two layers by the chain
rule~\eqref{eq:composed-jacobian}.  
From the user's perspective, the result
is a single differentiable function
$\Tmap : \haty \mapsto \ModelState$ that can be passed directly to any
gradient-based algorithm without further manual derivation.
\end{itemize}


The primary consequence of this decomposition is that extending the
sensitivity analysis to new parameterizations requires changes only in
the host code; the kernel primitive remains fixed.  
Conversely, improvements to the finite element solver or the DDM implementation
propagate automatically to every host-level workflow that uses the primitive.


The kernel developer implements the DDM once, exposing $\Mmap$ and its
differential rules as a fixed primitive.  
The end user writes the parameter map~$\Ymap$ in unrestricted host code and obtains end-to-end gradients automatically.  


The decomposition $\Tmap = \Mmap \circ \Ymap$ is not restricted to
static analysis or to any particular element type. 
Any analysis procedure that can be expressed as a differentiable primitive fits into the same pattern.  
Likewise, the parameter
map~$\Ymap$ can encode arbitrarily complex pre-processing logic which may include 
mesh generation, material homogenization, stochastic field sampling,
or any combination thereof.
