

\section{Literature}
% Hybrid Differentiation in Finite Element Analysis\label{hybrid-differentiation-in-finite-element-analysis}

\subsubsection{Direct Differentiation Method (DDM) in
FEA}\label{direct-differentiation-method-ddm-in-fea}

The \textbf{direct differentiation method (DDM)} computes sensitivities
by analytically differentiating the finite element governing equations
with respect to model parameters. In the earthquake engineering and
reliability community, DDM has been a preferred approach for obtaining
exact response gradients. 
For example, \cite{haukaas2003finite} implemented a comprehensive DDM-based sensitivity module in the OpenSees platform, enabling ``accurate, consistent and efficient computation of
response sensitivities'' for material, load, and even shape parameters.

In such implementations, one derives the partial derivative of each
element's stiffness matrix, internal force, etc., and assembles the
global gradient. 
The benefit is that the gradients are \textbf{exact (to
machine precision)} and often very efficient to evaluate. Studies have
shown that compared to finite difference or complex-step perturbation,
DDM yields sensitivities that are both \textbf{accurate and
computationally efficient}. 
\cite{scott2004resposne}, 
applied DDM to nonlinear beam-column elements and highlighted that it
offers a robust and efficient approach for response sensitivity analysis
of yielding structures.

Despite its advantages, DDM can be cumbersome to implement.
Every element formulation and material model in the FE code requires
hand-derived differentiation. 
This was feasible in research codes (e.g.~OpenSees) written in C++, but it poses a high barrier to integration with high-level scientific computing ecosystems (like Python). 
Maintaining and extending an analytically differentiated code
is labor-intensive and error-prone -- any change in the model requires
updating the derivatives. 
Thus, while DDM is well-known in structural optimization and reliability analysis for its efficiency, its use in
broader optimization or machine learning pipelines has been limited. 
In short, \textbf{DDM provides exact gradients at runtime with minimal
overhead, but demands significant upfront implementation effort}. 
This trade-off motivates exploring alternatives that can be more easily
generalized.


\subsubsection{Automatic Differentiation in Computational Mechanics}\label{automatic-differentiation-in-computational-mechanics}

In the past decade, \textbf{automatic differentiation (AD)} has made
significant inroads into computational mechanics, offering a way to
obtain exact gradients \emph{without} manual derivation. AD (also called
algorithmic differentiation or simply \emph{autodiff}) treats the FE
algorithm as a composition of differentiable operations and applies the
chain rule systematically. 

A thorough treatment of AD for iterative procedures such as the use of
the Newton-Raphson method for solving nonlinear equations is presented
by \textcite{beck1994automatic}, \textcite{gilbert1992automatic} and
\textcite{griewank1993derivative}. 
Historic accounts of the development
of AD can be found in \textcite{iri2009automatic}. 
Such treatments
generally begin with the work of Wengert (e.g. \textcite{wengert1964simple}). 
A more recent work,
\textcite{elliott2009beautiful} develops an elegant presentation and
implementation of forward-mode AD in the context of purely functional
programming.


Modern AD tools can compute Jacobians and
gradients \emph{programmatically}, and researchers have started
leveraging these tools in finite element contexts:

\begin{itemize}
\item
  \textbf{AD in FE Frameworks:} Large-scale multiphysics codes have
  begun to incorporate AD to form exact system Jacobians. A notable
  example is the MOOSE framework (INL) which uses the MetaPhysicL AD
  library. 
  \cite{lindsay2021automatic} describe the integration of AD
  into MOOSE's finite element routines, allowing the automatic
  computation of consistent tangent matrices for nonlinear solves. 
  They report that this approach yields an exact Jacobian at relatively small overhead, enhancing the robustness and convergence of Newton--Raphson solvers. 
  The key point is that developers can write the residual 
  equations in a natural way, and the AD system generates the Jacobian
  -- bypassing tedious manual differentiation.
\item
  \textbf{Differentiable FE Libraries:} Outside of traditional FE
  codebases, new libraries built on machine-learning frameworks have
  emerged. For example, \emph{Liang et al.} (2023) introduced
  \textbf{PyTorch-FEA}, which implements finite element analysis using
  PyTorch tensor operations. By ``taking advantage of autograd, an
  automatic differentiation mechanism in PyTorch,'' the library can
  seamlessly compute gradients of FE response with respect to input
  parameters. This enabled them to tackle inverse problems by coupling
  the FE model with neural networks. Notably, PyTorch-FEA was shown to
  \textbf{significantly speed up forward simulations} (benchmarking
  against Abaqus) while maintaining accuracy, and to improve inverse
  analysis accuracy and efficiency, especially when integrated with deep
  neural nets. The implication is that autodiff allows tight integration
  of FEM into optimization loops and machine learning pipelines --
  something that was traditionally cumbersome with low-level DDM codes.
\item
  \textbf{High-Performance Differentiable Simulation:} In tandem,
  efforts like \textbf{JAX-FEM} \citep{xue2023jaxfem} demonstrate
  that AD can be applied at scale. 
  JAX-FEM is a differentiable 3D FE solver built on Google's JAX library \citep{frostig2018compiling}, allowing GPU-accelerated computations. 
  Impressively, it handled a 3D problem with 7.7 million
  degrees of freedom and achieved an order-of-magnitude speedup over a
  commercial FE code by running on GPU. 
  JAX-FEM uses reverse-mode AD to
  obtain gradients, so one can perform \textbf{automatic inverse design}
  -- e.g.~topology optimization -- ``in a fully automatic manner
  \textbf{without} the need to manually derive sensitivities''. 
  This
  shows that with modern hardware and AD tools, even large-scale FE
  models can be made differentiable, opening the door to integrating FEM
  with AI-driven design. 
  (That said, JAX-FEM's success also relies on
  problem structure and substantial computing power; fully
  differentiating a million-DOF nonlinear problem is still extremely
  memory-intensive in general.)
\item
  \textbf{AD in Specialized Methods:} Researchers have also applied AD
  within specific computational mechanics techniques to simplify their
  implementation. 
  \cite{oberbichler2021efficient} used AD (with both forward and reverse modes) to derive the internal force vector and consistent stiffness matrix for
  \textbf{nonlinear isogeometric elements} from a single energy
  functional. 
  They specifically compared forward-mode vs.~reverse-mode
  AD for this task and found the reverse (adjoint) mode far more
  efficient for high-degree elements: the reverse-mode Jacobian required
  \emph{significantly fewer operations} and intermediate storage, and
  crucially, the number of AD intermediate variables no longer grew with
  the polynomial degree of shape functions. 
  This result underscores a
  general insight -- the choice of AD mode (forward or adjoint) must be
  tailored to the problem to avoid excessive overhead. In FE problems
  with many state variables but relatively fewer output quantities
  (e.g.~objective function or a few constraints), adjoint/reverse mode
  is usually optimal.
\end{itemize}

\textbf{Challenges:} While AD offers automation and exactness, it can
become \textbf{computationally prohibitive for entire nonlinear FE
simulations}. The memory footprint of reverse-mode AD is a known concern
-- e.g.~backpropagating through a long time-stepping simulation requires
storing or recomputing the state at each step. Some authors have noted
``strong limitations'' of naive AD when applied to general physics
problems, including high memory use and even issues with
non-differentiable points in simulations. 
In practice, fully
differentiable FE codes often require careful engineering
(checkpointing, code simplifications) to be feasible. This is where
\textbf{hybrid approaches} enter the picture -- combining analytic and
automatic differentiation to balance performance with flexibility.

\subsubsection{Hybrid Approaches}\label{hybrid-approaches-combining-ddm-and-ad}

Given the pros and cons of the two methods, a \textbf{hybrid
differentiation strategy} is attractive: use \emph{analytic
differentiation} for parts of the problem where we can efficiently
derive formulas, and use \emph{automatic differentiation} to handle the
rest. This idea has been explored in various forms.

% \textbf{Concept of Hybrid Differentiation:} 
\cite{bartlett2008hybrid} were early proponents of this approach in the context of
simulation codes. 
They observed that ``present AD tools\ldots rarely
compare to analytically derived versions'' in efficiency, but AD is
far easier to implement. 
Thus, \emph{``the focus of {[}their{]} work
{[}was{]} to combine the strength of these methods into a hybrid
strategy''}. In other words, one should \textbf{``select the
appropriate components''} of the simulation to differentiate manually
vs.~automatically to achieve an optimal balance of developer effort
and runtime cost. In their demonstration (fluid dynamics simulation in
C++), they hand-differentiated critical low-level routines and used
template-based AD for the rest -- yielding near-analytic performance
with much less manual coding. This philosophy directly inspires our
approach: we aim to differentiate the FE \emph{element kernel}
analytically (for efficiency), and let a higher-level AD system handle
glue code and complex couplings (for flexibility).

% \textbf{Contemporary Examples:} 
Very recently, \cite{dummer2024robust} put a hybrid approach into practice for nonlinear material modeling. 
They introduced a \emph{``semi-analytical split approach''}
  for a finite strain \textbf{generalized continuum} model. In their
  framework, the global finite element equations were derived
  analytically as usual, but the \textbf{material's stress-update
  algorithm} (which was quite complex, involving a gradient-enhanced
  damage model) was differentiated using automatic differentiation
  \textbf{with hyper-dual numbers}. Hyper-dual AD provided first and
  second derivatives of the stress update in a black-box fashion,
  yielding an exact consistent tangent matrix for the material
  sub-problem. This \textbf{split} saved them from manually
  differentiating an arduous constitutive law, while still embedding
  that law into an efficient FE solver. 
  In effect, Dummer \emph{et al.} achieved a modular differentiation: the element
  stiffness contributions from a complicated material are obtained via
  AD, and everything else (element formulation, assembly, solver) uses
  standard analytic differentiation.

% \textbf{Adjoint-AD Hybrids:} 
Another flavor of hybridization is mixing \emph{adjoint methods} with AD. 
Adjoint (reverse) differentiation is itself an analytic approach at the system level, often used in topology optimization or sensitivity analysis to avoid recomputing a full Jacobian. 
One can use an adjoint equation to get the gradient of a single objective with respect to many inputs very efficiently -- but one still needs partial derivatives (Jacobian blocks) inside that adjoint equation. 
%
\cite{oberbichler2021efficient} provide a nice
example: they apply the adjoint method to the \textbf{global} IGA
system (yielding what they call a ``core-congruential formulation''
separating geometry from mechanics), but they obtain the necessary
\textbf{element-level derivatives} via algorithmic differentiation of
the energy functional. 
The result was a \textbf{clean separation}: the
CAD-based geometry routines were automatically differentiated, and
then the adjoint method assembled the total sensitivity with far fewer
operations than a pure forward differentiation would require. This
hybrid of AD and adjoint analytic methods allowed them to tackle shape
sensitivities for very high-order elements efficiently. We anticipate
a similar layering in our work: the low-level FE code provides
\emph{analytic} derivatives (DDM-style) for element responses, which
can then be used in high-level autodiff (which functions much like an
adjoint when computing gradients of a larger objective that includes
the FE model).

In summary, the literature suggests that \textbf{combining direct and
automatic differentiation is a promising route} to bring finite elements
into the realm of optimization and machine learning. 
Pure AD-based FEM implementations exist and have shown what's possible (e.g.~end-to-end differentiable simulators like JAX-FEM and PyTorch-FEA). However, they also highlight the downsides -- namely, the runtime and memory costs -- when applied to large, nonlinear FE problems. 
On the other hand, decades of work on DDM in computational mechanics have given us methods to get
gradients efficiently for very specific problems (e.g.~the OpenSees DDM
for structural sensitivities, or recent DDM for geometrically-nonlinear
frame elements). 

The \textbf{hybrid approach seeks to get the ``best of
both worlds.''} By embedding an analytically differentiated FE
\emph{kernel} inside a higher-level AD framework, we retain the
efficiency of hand-coded derivatives for the heavy lifting (the finite
element internal calculations) and the flexibility of automatic
differentiation for coupling with other software (optimization
algorithms, machine learning models, etc.). 
This approach has only started to appear in the literature in the last few years, and to our knowledge, no general-purpose Python-based FE tool with built-in DDM
sensitivities exists yet. 
Our work will fill that gap. 
We will demonstrate how a \textbf{DDM-enhanced FE solver} can be \emph{embedded}
in an autodiff environment to enable gradient-based design,
optimization, and learning on complex FE models -- \emph{without} the
prohibitive memory cost of differentiating through the entire nonlinear
solve. This aligns with the trends seen in recent research and pushes
them a step further by uniting the efficiency of classical methods with
the flexibility of modern autodiff frameworks.




