\subsection{Parameter binding in the evaluation trace}
\label{sec:parameter-binding}

The differential formulas of Section~\ref{sec:differential} assume
that the coordinates of \(\CoreParameters\) are already identified.
In a model-building workflow this identification is itself part of the
construction. A host variable is introduced during the evaluation of a
model factory, is attached to one or more intermediate objects, and
ultimately acts on one or more concrete parameter locations in the
finite element kernel. The purpose of the present subsection is to
describe this binding process in terms that are closer to the
literature on automatic differentiation.

\paragraph{Evaluation trace.}
Let
\[
\HostParameters = (\hat y_1,\ldots,\hat y_{n_{\mathrm h}})
\]
denote the scalar host coordinates. The model factory is evaluated as
a finite sequence of elementary construction operations. 
This sequence may be regarded as an evaluation trace, or equivalently as a
computational graph whose nodes are the intermediate objects created during model construction.

Let
\[
\mathcal O
=
\mathcal O_{\mathrm{mat}}
\cup
\mathcal O_{\mathrm{sec}}
\cup
\mathcal O_{\mathrm{fib}}
\cup
\mathcal O_{\mathrm{ele}}
\]
denote the set of model objects appearing in this trace, partitioned
into materials, sections, fibers, and elements. The trace induces a
directed dependency relation on \(\mathcal O\). For example, a fiber
depends on the section in which it is created, and an element depends
on the section instances referenced in its definition.

\paragraph{Labelled coordinates.}
Each scalar host coordinate is treated together with an identifying
label,
\[
\pi_i = (i,\nu_i,\hat y_i),
\qquad i=1,\ldots,n_{\mathrm h},
\]
where \(i\) is a global differentiation index and \(\nu_i\) is a
symbolic name. The value \(\hat y_i\) is used in the primal
construction, while \((i,\nu_i)\) records the identity of the
coordinate in the differential structure.

\paragraph{Binding.}
When a labelled coordinate \(\pi_i\) appears in the evaluation trace,
it is bound to an object
\[
a_i \in \mathcal O,
\]
called its anchor. The anchor specifies the level at which the
coordinate enters the model description. A modulus may be anchored at
a material object, a geometric parameter at a section object, and a
fiber area at a fiber object.

The evaluation trace then propagates this dependence downstream
through the computational graph. In this way, one obtains the set of
all concrete parameter locations in the kernel that are controlled by
coordinate \(i\).

\paragraph{Kernel addresses.}
Let \(\mathcal A\) denote the set of admissible kernel addresses. An
address \(\alpha \in \mathcal A\) identifies a concrete parameter slot
in the assembled model, for example a material parameter at a given
section point within a given element, or a fiber attribute within a
section occurrence.

The binding process determines a set-valued map
\[
\mathcal B : \{1,\ldots,n_{\mathrm h}\} \to 2^{\mathcal A},
\]
where \(\mathcal B(i)\) is the set of all kernel addresses reached by
following the dependency graph outward from the anchor \(a_i\).
Accordingly, \(\mathcal B(i)\) is the support of the global coordinate
\(i\) in the assembled model.

\paragraph{Meaning of a core coordinate.}
Once the binding map \(\mathcal B\) has been resolved, the host
coordinates acquire a concrete realization in the kernel. In the
simplest case one identifies
\[
\CoreParameters = (y_1,\ldots,y_{n_{\mathrm c}}),
\qquad
n_{\mathrm c}=n_{\mathrm h},
\qquad
y_i=\hat y_i,
\]
so that the host and kernel use the same scalar coordinates, but the
binding map \(\mathcal B\) specifies where each coordinate acts in the
assembled model.

With this interpretation, the derivative
\[
\frac{\partial \ResponseMap_{\ModelFamily}}{\partial y_i}
\]
means the sensitivity obtained by perturbing simultaneously every
kernel address in \(\mathcal B(i)\), while holding all remaining
coordinates fixed. Thus the kernel derivative is not attached to a
single local slot, but to the full support of a global coordinate
throughout the evaluation trace.

\paragraph{Forward contraction.}
The forward differential therefore takes the form
\[
\D \ResponseMap_{\ModelFamily}(\CoreParameters)[\dot{\CoreParameters}]
=
\sum_{i=1}^{n_{\mathrm c}}
\frac{\partial \ResponseMap_{\ModelFamily}}{\partial y_i}(\CoreParameters)\,
\dot y_i,
\]
where each partial derivative already accounts for all bound kernel
addresses associated with coordinate \(i\). The host layer determines
the perturbation coefficients \(\dot y_i\), while the kernel layer
evaluates the corresponding directional effect on the response.

\paragraph{Role in the overall framework.}
The binding map does not alter the chain rule of
Section~\ref{sec:differential}. Its role is more basic. It identifies
the coordinates with respect to which the kernel derivative is to be
understood. In automatic-differentiation language, it supplies the
missing connection between the host-side evaluation trace and the
kernel-side response map by recording how labelled host coordinates
are bound to concrete parameter addresses in the assembled finite
element model.

\begin{remark}
The preceding description assumes that the evaluation trace produces a
fixed model topology. Under this assumption the binding map
\(\mathcal B\) is locally constant, and the composite response is
differentiable in the usual sense. If the topology itself changes with
the host parameters, then the map \(\ParameterMap\) is only piecewise
smooth, and the resulting differential structure requires additional
care.
\end{remark}